A comparison of single-cell trajectory inference methods: towards more accurate and robust tools

### 2018-10-02 manuscript/paper.Rmd
update paper

While not directly related to the accuracy of the inferred trajectory, the quality of the implementation is also an important evaluation metric [@_doesyourcode_2018]. To assess this, we scored each method using a transparent checklist of important scientific and software development practices, including software packaging, documentation, automated code testing, and publication into a peer-reviewed journal (`r ref("stable", "qc_scorsheet")`).

@@

We ran each method on each dataset for maximally 1 hour and gave each process 10GB of memory. To determine the running time of each method, we started the timer right after data loading, and the loading of any packages, and stopped the clock before post-processing and saving of the output. Pre- and postprocessing steps specific to a method, such as dimensionality reduction and gene filtering, were included in the time. To estimate the maximal memory usage, we used the max_vmem value from the qacct command provided by the gridengine cluster. We acknowledge that these values are very noisy, and the averages provided in this study are therefore only rough estimates.

The relationship between the dimensions of a dataset and the running time or maximal memory usage was modelled using generalized additive models. We used thin-plate regression spline as smoothing basis, with $log_{10}|cells|$ and $log_{10}|features|$ as predictor variables, and fitted this model using the gam as implemented in the R mgcv package, with $log_{10}\textit{time}$ (or $log_{10}\textit{memory}$) as outcome.

>>>>

While not directly related to the accuracy of the inferred trajectory, the quality of the implementation is also an important evaluation metric [@_doesyourcode_2018]. To assess this, we scored each method using a transparent checklist of important scientific and software development practices, including software packaging, documentation, automated code testing, and publication into a peer-reviewed journal (`r ref("stable", "qc_scorseheet")`).

@@

We ran each method on each dataset for maximally 1 hour and gave each process 10GB of memory. To determine the running time of each method, we started the timer right after data loading and the loading of any packages, and stopped the clock before post-processing and saving of the output. Pre- and postprocessing steps specific to a method, such as dimensionality reduction and gene filtering, were included in the time. To estimate the maximal memory usage, we used the max_vmem value from the qacct command provided by the gridengine cluster. We acknowledge however that these values are very noisy, and the averages provided in this study are therefore only rough estimates.

The relationship between the dimensions of a dataset and the running time or maximal memory usage was modelled using generalized additive models. We used thin-plate regression spline as smoothing basis, with $log_{10}|\textit{cells}|$ and $log_{10}|\textit{features}|$ as predictor variables, and fitted this model using the gam as implemented in the R mgcv package, with $log_{10}\textit{time}$ (or $log_{10}\textit{memory}$) as outcome.
